Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{ Funded by DARPA Cyber Assured Systems Engineering (CASE) contract HR00111890001. }

### 2021-05-07 paper/example.tex
Add minor revisions to example

\figref{fig:example} is an AADL description of an implementation of a software system for route planning and automated control for a UAV. It is loosely based on the system in the case study presented later. The system receives an automation request that is forwarded to an untrusted third-party route planner (AI) that decides the flight path of the UAV based on its current position and the requested task. The waypoint manager (WM) receives the mission command as a set of waypoints from the planner and starts the UAV flying the mission, issuing waypoints to the UAV flight controller as the UAV location changes. It is an \emph{as is} legacy component. The alert manager never raises the alert as there is no capacity in the design to know when an alert is needed.

@@

The contract in \figref{fig:sw} uses \texttt{eq} statements to define variables local to the contract. For example, the \texttt{req} variable is equivalent to the \texttt{event(req)} expression. There is an implicit \emph{event} input (or output) associated with every named event port in a component. The semantics used here do not not buffer events so the implicit input (or output) is a boolean value. An \texttt{event} expression refers to that implicit input (or output) and is true if there is data on the named port. The system contract assumes well-formed input and guarantees a set of properties about the output.

@@

The guarantee is an invariant on the expression \texttt{policy or since} meaning that either the policy holds or the alert is sounding. The \texttt{policy} is satisfied if there is a request with a response or nothing in the current time step (\texttt{today}) or a request yesterday with its response today. The value of \texttt{yesterday} in the current time step relies on values from the previous time step. The \texttt{yesterday} value uses the \texttt{->} operator for initialization. The left operand is the initial value of \texttt{yesterday} at the start of the system, which in this example is \texttt{(req and not rsp)} because seeing a request with no response is inconclusive at the start.  The right operand is the value of \texttt{yesterday} after the initial step. Here the \texttt{pre} operator refers to the value of the expression \texttt{(req and not rsp)} in the prior time step. Intuitively, \texttt{yesterday} is true if the previous time step made a request without a matching response and the current time step has the matching response to that request with no new request.

The value of \texttt{since} relies on its own value in the previous time step. The intuitive reading of the expression is that the alert has been true since the time when it first sounded. The first \texttt{alrt} sets \texttt{since} to true, and once the value of \texttt{since} is true, that value persists as long as \texttt{alrt} holds. The \emph{Alert if start is not bounded relative to a request} guarantee defines one behavior of a correctly implemented system. Together with the other guarantees, the contract models the expected input and output of the system as a whole. Contracts also model the expected behavior of each component used to implement the system.

Model checking proves whether the contract models of the components, with their connections, implement the system model in \figref{fig:sw}. The results of the model checker are in \figref{fig:example}(b). The red exclamation points are properties that do not hold. Each comes with a counter-example. The result is not unexpected since the model of the route planner leaves the output behavior unconstrained (i.e., an untrusted component can do anything). The counter-example showing each vio with a counterexample that violates the policy and does not raise the alert.

@@

A monitor captures a relation over time on input data and is able to reason about temporal properties of that input. By definition, it raises an alert output when the relation is violated. The generated monitor contract is shown in \figref{fig:assurance}(b). The \texttt{is\_latched} value makes the alert persistent meaning that once raised always raised. The definition for \texttt{policy} is taken from the contract in \figref{fig:sw} since this high-assurance component is the monitor that needs to alert violations of the bounded response policy. As before the guarantees for the outputs are autogenerated by the tool.

>>>>

\figref{fig:example} is an AADL description of an implementation of a software system for route planning and automated control for a UAV. It is loosely based on the system in the case study introduced later. The system receives an automation request that is forwarded to an untrusted third-party route planner (AI) that decides the flight path of the UAV based on its current position and the requested task. The waypoint manager (WM) receives the mission command as a set of waypoints from the planner and starts the UAV flying the mission, issuing waypoints to the UAV flight controller as the UAV location changes. It is an \emph{as is} legacy component. The alert manager never raises the alert as there is no capacity in the design to know when an alert is needed.

@@

The contract in \figref{fig:sw} uses \texttt{eq} statements to define variables local to the contract. For example, the \texttt{req} variable is equivalent to the \texttt{event(AutomationRequest)} expression. There is an implicit \emph{event} input (or output) associated with every named event port in a component. The semantics used here do not not buffer these events so the implicit input (or output) is only a boolean value. An \texttt{event} expression refers to that implicit input (or output) and is true if there is data on the named port. The system contract assumes well-formed input and guarantees a set of properties about the output.

@@

The guarantee is an invariant on the expression \texttt{policy or since} meaning that either the policy holds or the alert is sounding. The \texttt{policy} is defined by two local values: \texttt{today} and \texttt{yesterday}. The \texttt{today} value is asserted when in the current time step there is a request with a response, or there is no request and no response.  The value of \texttt{yesterday} in the current time step relies on values from the previous time step \texttt{->} operator for initialization as the previous time step is undefined in the first step of the system. The left operand is the initial value of \texttt{yesterday} at start, which in this example is \texttt{(req and not rsp)}, because seeing a request with no response is inconclusive in the first step of the system. The right operand is the value of \texttt{yesterday} after the initial step. Here the \texttt{pre} operator refers to the value of the expression \texttt{(req and not rsp)} in the prior time step. Intuitively, \texttt{yesterday} is true if the previous time step made a request without a matching response and the current time step has the matching response to that request with no new request.

The value of \texttt{since} in \emph{Alert if start is not bounded relative to a request} relies on its own value in the previous time step. The intuitive reading of the expression is that the alert has been true since the time when it first sounded. The first \texttt{alrt} sets \texttt{since} to true, and once the value of \texttt{since} is true, that value persists as long as \texttt{alrt} holds. The \emph{Alert if start is not bounded relative to a request} guarantee defines one behavior of a correctly implemented system. Together with the other guarantees, the contract models the expected input and output of the system as a whole.

Model checking proves whether the contract models of the components, with their connections, implement the system model in \figref{fig:sw}. The results of the model checker are in \figref{fig:example}(b). The red exclamation points are properties that do not hold. Each comes with a counter-example. The result is not unexpected since the model of the route planner leaves the output behavior unconstrained (i.e., an untrusted component can do anything). The counter-example for \emph{Alert if start is not bounded relative to a request} shows a response in the first time step with no matching request.

@@

A monitor captures a relation over time on input data and is able to reason about temporal properties of that input. By definition, it raises an alert output when the relation is violated. The generated monitor contract is shown in \figref{fig:assurance}(b). The \texttt{is\_latched} value makes the alert persistent meaning that once raised always raised. The definition for \texttt{policy} is taken from the contract in \figref{fig:sw}. As before the guarantees for the outputs are autogenerated by the tool.
